### 2023-06-01 paper/ICSEfixnotes.tex
icse fix core

1. emphasize that comby generated way more mutants in every case than standard tools (I can't see how this is true, but somehow the reviewers had a big discussion and concluded the opposite???)

2.  we genuinely should take the time to improve the equivalence estimates, and try to do PIT

3.  go into more detail on supported operators early on, list numbers of rules per universal, C-like, per-language implementation, and discuss a) which operators we cannot currently support and b) how other tools that in theory could support those also do not support them, and that e.g., variable replacement is present in early literature but in practice has been part of almost no serious mutation implementations