### 2023-06-01 paper/ICSEfixnotes.tex
more detail

1. emphasize that comby generated way more mutants in every case than standard tools (I can't see how this is true, but somehow the reviewers had a big discussion and concluded the opposite???)

2.  we genuinely should take the time to improve the equivalence estimates, and try to do PIT (my proposal:  do up to 30 unkilled mutants, and look at PIT source reports to try to figure equivalency there)

3.  go into more detail on supported operators early on, list numbers of rules per universal, C-like, per-language implementation, and discuss a) which operators we cannot currently support and b) how other tools that in theory could support those also do not support them, and that e.g., variable replacement is present in early literature but in practice has been part of almost no serious mutation implementations

4. change way we cite the ICSE 2018 tool paper; redact the details and flat say we had an early tool paper on the regexp idea with little data but it's ours, we are the same tool